Supplementary Information for\\ ``A weakly structured stem for human origins in Africa''

### 2023-03-15 supplement.tex
minor edits
question

Thus, either an incipient OOA population must have expanded beyond the Near
East by 85ka and then given rise to today’s Eurasian populations or a
successful expansion would likely post-date 65ka. Morphological and
archaeological evidence for Homo sapiens in the Near East reappears by 50ka in
the Levant with the Initial Upper Paleolithic stone tool industry
\citep{Boaretto2021-oi} and a cranium from Manot Cave
\citep{Hershkovitz2015-lq}. There is also strong consensus that a widespread
human presence in Europe and Asia is dated to after 45ka. Archaeological
\citep[e.g.,][]{Fewlass2020-se}, morphological \citep[e.g.,][]{Hublin2020-nk},
and genetic data \citep[e.g.,][]{Prufer2021-ao} all indicate that about 45ka
\emph{Homo sapiens} had expanded into eastern Europe and within 5,000 years
rapidly colonized the entire span of southern Europe. This period further
coincides with the dates for the last Neanderthal presence in Europe at
approximately 40ka \citep{Deviese2017-de,Hublin2017-uy}. Similarly in Asia,
evidence of the Initial Upper Paleolithic occurs 39-45ka as far east as Lake
Baikal. Characteristics of the IUP support a coherent, cultural package which
may have been transmitted by demic expansion \citep{Kuhn2014-kj,Zwyns2021-zv}.
By 45ka, there is additionally an unequivocal \emph{H. sapiens} femur from
southern Siberia which contains derived alleles shared with later Eurasian
populations \citep{Fu2014-am}. The Ust’-Ishim
specimen contains evidence of admixture with Neanderthals, at a level similar
to contemporary East Asian genomes. This supports estimates of Neanderthal
admixture occurring earlier than 45ka, estimated to be by 47-65ka as inferred
from Neanderthal tract lengths \citep{Sankararaman2012-uw}.

>>>>

Thus, either an incipient OOA population must have expanded beyond the Near
East by 85ka and then given rise to today’s Eurasian populations or a
successful expansion would likely post-date 65ka. Morphological and
archaeological evidence for Homo sapiens in the Near East reappears by 50ka in
the Levant with the Initial Upper Paleolithic stone tool industry
\citep{Boaretto2021-oi} and a cranium from Manot Cave
\citep{Hershkovitz2015-lq}. There is also strong consensus that a widespread
human presence in Europe and Asia is dated to after 45ka. Archaeological
\citep[e.g.,][]{Fewlass2020-se}, morphological \citep[e.g.,][]{Hublin2020-nk},
and genetic data \citep[e.g.,][]{Prufer2021-ao} all indicate that about 45ka
\emph{Homo sapiens} had expanded into eastern Europe and within 5,000 years
rapidly colonized the entire span of southern Europe. This period further
coincides with the dates for the last Neanderthal presence in Europe at
approximately 40ka \citep{Deviese2017-de,Hublin2017-uy}. Similarly in Asia,
evidence of the Initial Upper Paleolithic occurs 39-45ka as far east as Lake
Baikal. Characteristics of the IUP \sgcomment{define?} support a coherent, cultural package which
may have been transmitted by demic expansion \citep{Kuhn2014-kj,Zwyns2021-zv}.
By 45ka, there is additionally an unequivocal \emph{H. sapiens} femur from
southern Siberia which contains derived alleles shared with later Eurasian
populations \citep{Fu2014-am}. The Ust’-Ishim
specimen contains evidence of admixture with Neanderthals, at a level similar
to contemporary East Asian genomes. This supports estimates of Neanderthal
admixture occurring earlier than 45ka, estimated to be by 47-65ka as inferred
from Neanderthal tract lengths \citep{Sankararaman2012-uw}.
